\section{Functional requirements}
\subsection{Video input and preprocessing}
\begin{itemize}
    \item The system accepts video inputs via URL (e.g., YouTube links) or direct file upload (supported formats: MP4, AVI, MOV; max file size 500MB). 
    \item The system transcribes audio to text using Whisper, handling accents, background noise, and multiple speakers with at least 85\% accuracy on clear audio. 
    \item Timestamps are preserved in the transcription for linking back to video segments. 
    \item Error handling: If transcription fails (e.g., due to unsupported format), provide user-friendly error messages and retry options.
\end{itemize}

\subsection{Summarization module}
\begin{itemize}
    \item The system generates extractive or abstractive summaries from the transcribed text, condensing content to 10-20\% of original length. 
    \item Key features: Extract main topics, keypoints, and conclusions using NLP models ; support keyword highlighting and mind-map generation as an extension. 
    \item Output formats: Plain text, bullet points, or PDF export with timestamps for navigation. 
    \item Personalization: Allow users to specify focus areas (e.g., "summarize technical terms only") or summary style (e.g., concise vs. detailed).
    \item Integration: Link summary sections to quiz questions for cross-referencing.
\end{itemize}

\subsection{Quiz generation module}
\begin{itemize}
    \item The system automatically generates quizzes based on the summary and transcription, including 5-15 questions per video (user-adjustable).
    \item Supported quiz types: Multiple-choice (with 4 options, one correct), short-answer, true/false, and fill-in-the-blank. 
    \item Each question shall include: Correct answer, explanation (derived from video content).
    \item Features: Generate adaptive quizzes (e.g., harder questions on weak topics based on user performance).
    \item Output: Interactive quiz interface (e.g., via web app) with scoring, feedback, and export to formats like Google Forms or PDF.
\end{itemize}

\subsection{User Interface and Interaction}
\begin{itemize}
    \item The system provides a web-based UI for input submission, progress tracking,  and result viewing. 
    \item User authentication: Support guest access for basic use; optional login for saving quizzes or history
    \item Interactivity: Allow users to edit generated summaries/quizzes (e.g., regenerate specific questions) and provide feedback for model improvement.
    \item Reporting: Generate usage logs or analytics (e.g., average quiz score) for educators.
\end{itemize}